\begin{acknowledgments}
First and foremost, I would like to thank my advisor, Professor Umakishore Ramachandran, for his continuous support and motivation during my path as a PhD student. I routinely claim to be one of the luckiest PhD students ever to have an advisor like Prof. Kishore. He has helped me constantly both by steering my research efforts in the right direction and by helping me maneuver out of technical roadblocks. More than that, his kind and understanding attitude toward his students has made my PhD journey much more enjoyable than I would have imagined. I will forever be grateful and indebted to him for the support he has provided to me.
\par I would like to thank the members of my dissertation committee for taking their valuable time to be on my committee and providing feedback that has significantly improved my dissertation. I would specifically like to thank Professor Ada Gavrilovska for mentoring me in my earlier years as a PhD, when we worked on a project together. She has always encouraged me to aim higher in my academic journey. Next, I would like to thank Prof. Alexandros Daglis, for his support and contribution toward the OneEdge project. His inputs helped us significantly improve the paper and get it accepted for publication. I would like to thank Dr. Abhigyan Sharma for his continued encouragement and help that he has provided ever since I have known him, especially his guidance for the job search. Finally, I would like to thank Prof. Ahmed Saeed for helping me improve my dissertation during my proposal.
\par As with any journey, fellow travelers are crucial to make it a good experience. I express a special thanks to my colleague Enrique Saurez, with whom I have been working from day 1 of my PhD to almost the very end. Not only have we collaborated extensively for research, we have also shared a lot of good times outside school. He has also selflessly helped me out on a number of occasions and made life a lot easier for me. In the same vein, I would like to thank my colleagues Adam Hall and Zhuangdi Xu, for not only collaborating me and helping me with my dissertation, but also being there as close friends for almost the entirety of my stay at Georgia Tech. I am grateful for all my wonderful colleagues at the Embedded Pervasive Laboratory (EPL) -- Tyler Landle, Anirudh Sarma, Manasvini Sethuraman, Alan Nussbaum, Jinsun Yoo, Difei Cao, Ashish Bijlani and Wonhee Cho. I am also grateful to my collaborators from the University of Stuttgart -- Prof. Kurt Rothermel, Ruben Mayer, Henriette Roger, and Sukanya Bhowmik, for the interesting work we did with each other. I am also indebted to my external collaborator from AT\&T Alex Zelezniak for lending his immense expertise to my projects.
\par On a personal note, I thankful beyond words for my family -- my mother and sister and my late father -- for their never ending love and encouragement on every step of the way. I am indebted to my very dear friend Nivya Sharma, who has been a part of my life for the past year and a half and helped me maneuver through the most difficult and hectic part of my PhD. Last but not least, I would like to thank my friends Ipsita and Shubhayan who have helped me throughput my stay at Georgia Tech with anything I needed.
\end{acknowledgments}
